Inference for Trans-dimensional Bayesian Models with Diffusive Nested Sampling

### 2014-12-06 Documents/Paper/ms.tex
Talking about the trace plot

If we were to do this analysis by trying to explore the posterior directly,
rather than by using Nested Sampling, it would be difficult to jump between
the $N=1$ solution and the $N=2$ solution, because the posterior distribution
would be composed of a mixture of a small-volume but high-likelihood spike
and a large-volume but (relatively) low likelihood
slab. Note that this phase transition still exists if we condition on $N=2$,
so is not caused by the trans-dimensional model.

>>>>

If we were to do this analysis by trying to explore the posterior directly,
rather than by using Nested Sampling, it would be difficult to jump between
the $N=1$ solution and the $N=2$ solution, because the posterior distribution
would be composed of a mixture of a small-volume but high-likelihood spike
and a large-volume but (relatively) low likelihood
slab. Note that this phase transition still exists if we condition on $N=2$,
so is not caused by the trans-dimensional model. To demonstrate this,
Figure~\ref{fig:trace_logl} shows the value of the log likelihood versus time
for 45 million iterations of MCMC targeting the posterior distribution.
The sampler should visit states with log likelihood $\sim -745$ about 10\%
of the time (corresponding with models that don't contain the small oscillation),
and the rest of the time should be spent in states with log likelihood around
$-730$. In this run transitions between these two phases occurred only a
handful of times. In a practical application, computing the posterior probability
for the existence of the small oscillation might be the whole point of the
calculation, so mixing between models where it is present and models where it
is absent is crucial.
